---
title: "Data Dictionary"
author: "Rob Kemp"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Dictionary}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---
```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(codemog)
knitr::opts_chunk$set(
  comment = "#>",
  error = FALSE,
  tidy = FALSE)
```
The `codemog` package contains population estimates and forecasts generated by the Colorado State Demography Office (located within the Department of Local Affairs).  There are currently 4 datasets available through this R Package.  The following is a detailed dictionary of each dataset.

### county_est: Colorado County Population Estimates 

Each year the State Demography Office creates population estimates for all of the counties in Colorado.  The Estimates process requires a variety of input data: births and deaths from the Department of Public Health and Environment, Net Migration from the U.S. Census Bureau, as well as Construction and Group Quarters data from local governments.  This process generates more data than just a total population estimate.

This is a snapshot of the data:
```{r}
county_est
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions
#####geonum: 
numeric, for merging with data from SDO PostgreSQL Census databases.

#####countyfips: 
Federal Information Processing Standards (FIPS) Number for each County, without leading zeros.

#####county: 
Name of the County

#####year: 
Estimate year, specifically reflects July 1 of each year.

#####totalPopulation: 
Total Population on July 1 of the year.

#####householdPopulation: 
Population Living in Households on July 1 of the year.

#####groupquartersPopulation: 
Population living in group quarters (e.g. nursing homes, prisons, dorms) on July 1 of the year.

#####totalHousingUnits: 
Estimates number of housing units for July 1 of the year.

#####occupiedHousingUnits: 
Estimates of the number of occupied housing units on July 1, this is also a measure of the number of households.

### county_forecast: Colorado Municipal Population Estimates
Each year the State Demogrpahy Office creates a population forecast and historical estimate for each county by age and sex.  This process uses a cohort component model matched with an economic model to link jobs estimates with population estimates.  The model creates estimates for years that the Estimates program creates estimates for and forecasts for the years from then unitl 2040.  

This is a snapshot of the data:
```{r}
county_forecast
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions

#####countyfips: 
Federal Information Processing Standards (FIPS) Number for each County, without leading zeros.


#####year: 
Estimate year, specifically reflects July 1 of each year.

#####age:
Single year of age, topcoded at 90.  A value of 90 contains the population for all those that are 90 and above.  

#####county: 
Name of the County

#####malePopulation: 
Total male population on July 1 of the year.

#####femalePopulation: 
Total female population on July 1 of the year.

#####totalPopulation: 
Total population on July 1 of the year.

#####datatype:
This takes the value of "Estimate" when the data is an estimate and "Forecast" when the data is a forecast.

### muni_est: Colorado Municipal Population Estimates 

Each year the State Demography Office creates population estimates for all of the municipalities in Colorado.  The Estimates process requires Construction and Group Quarters data from local governments and Census estimates of persons per household and a base year population and housing unit estimate.  This process generates more data than just a total population estimate.

This is a snapshot of the data:
```{r}
muni_est
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions
#####geonum: 
numeric, for merging with data from SDO PostgreSQL Census databases.

#####placefips: 
Federal Information Processing Standards (FIPS) Number for each municipality, without leading zeros.

#####municipality: 
Name of the Municipality

#####year: 
Estimate year, specifically reflects July 1 of each year.

#####totalPopulation: 
Total Population on July 1 of the year.

#####householdPopulation: 
Population Living in Households on July 1 of the year.

#####groupquartersPopulation: 
Population living in group quarters (e.g. nursing homes, prisons, dorms) on July 1 of the year.

#####totalHousingUnits: 
Estimates number of housing units for July 1 of the year.

#####occupiedHousingUnits: 
Estimates of the number of occupied housing units on July 1, this is also a measure of the number of households.

### muni_win_est: Colorado Municipal Population Estimates 

Each year the State Demography Office creates population estimates for all of the municipalities in Colorado.  The Estimates process requires Construction and Group Quarters data from local governments and Census estimates of persons per household and a base year population and housing unit estimate.  This process generates more data than just a total population estimate. Municipalities do not always exist entirely within one county, so parts exist in each county.  These parts are an additional feature of this dataset (i.e. you can get specific estimates for the Adams, Arapahoe, and Douglas County parts of Aurora).

This is a snapshot of the data:
```{r}
muni_win_est
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions
#####geonum: 
numeric, for merging with data from SDO PostgreSQL Census databases.

#####countyfips: 
Federal Information Processing Standards (FIPS) Number for each County, without leading zeros.

#####placefips: 
Federal Information Processing Standards (FIPS) Number for each municipality, without leading zeros.

#####municipality: 
Name of the Municipality

#####year: 
Estimate year, specifically reflects July 1 of each year.

#####totalPopulation: 
Total Population on July 1 of the year.

#####householdPopulation: 
Population Living in Households on July 1 of the year.

#####groupquartersPopulation: 
Population living in group quarters (e.g. nursing homes, prisons, dorms) on July 1 of the year.

#####totalHousingUnits: 
Estimates number of housing units for July 1 of the year.

#####occupiedHousingUnits: 
Estimates of the number of occupied housing units on July 1, this is also a measure of the number of households.


 ### county_hist: Colorado County Historical Population Estimates 

Each year the State Demography Office creates population estimates for all of the counties in Colorado.  The Estimates process requires a variety of input data: births and deaths from the Department of Public Health and Environment, Net Migration from the U.S. Census Bureau, as well as Construction and Group Quarters data from local governments.  This data includes estimates up until 2009, the year prior to our current estimates base. 

This is a snapshot of the data:
```{r}
county_hist
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions
#####geonum: 
numeric, for merging with data from SDO PostgreSQL Census databases.

#####countyfips: 
Federal Information Processing Standards (FIPS) Number for each County, without leading zeros.

#####county: 
Name of the County

#####year: 
Estimate year, specifically reflects July 1 of each year.

#####totalPopulation: 
Total Population on July 1 of the year.

#####datatype
Describes whether data is an estimate or a forecast.

### muni_hist: Colorado Municipal Historical Population Estimates 

Each year the State Demography Office creates population estimates for all of the municipalities in Colorado.  The Estimates process requires Construction and Group Quarters data from local governments and Census estimates of persons per household and a base year population and housing unit estimate.  This process generates more data than just a total population estimate.

This is a snapshot of the data:
```{r}
muni_hist
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions
#####geonum: 
numeric, for merging with data from SDO PostgreSQL Census databases.

#####placefips: 
Federal Information Processing Standards (FIPS) Number for each municipality, without leading zeros.

#####municipality: 
Name of the Municipality

#####year: 
Estimate year, specifically reflects July 1 of each year.

#####totalPopulation: 
Total Population on July 1 of the year.


### muni_win_hist: Colorado Municipal Historical Population Estimates 

Each year the State Demography Office creates population estimates for all of the municipalities in Colorado.  The Estimates process requires Construction and Group Quarters data from local governments and Census estimates of persons per household and a base year population and housing unit estimate.  This process generates more data than just a total population estimate. Municipalities do not always exist entirely within one county, so parts exist in each county.  These parts are an additional feature of this dataset (i.e. you can get specific estimates for the Adams, Arapahoe, and Douglas County parts of Aurora).

This is a snapshot of the data:
```{r}
muni_win_hist
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions
#####geonum: 
numeric, for merging with data from SDO PostgreSQL Census databases.

#####countyfips: 
Federal Information Processing Standards (FIPS) Number for each County, without leading zeros.

#####placefips: 
Federal Information Processing Standards (FIPS) Number for each municipality, without leading zeros.

#####municipality: 
Name of the Municipality

#####year: 
Estimate year, specifically reflects July 1 of each year.

#####totalPopulation: 
Total Population on July 1 of the year.

 ### county_jobs: Colorado County Jobs Estimates (SDO Version)

Each year the State Demography Office creates total jobs estimates by sector.  These jobs estimates include wage and salary employment, sole proprietors, and agricultural employment estimates.  These data are made each year and span from 2001 to 2013.  Due to supression requirements from the BLS, some industries are noted with an S for supressed.  

This is a snapshot of the data:
```{r}
county_jobs
```
NOTE: The dataframe is tagged as a `tbl_df()` class.  If you use `dplyr` then this will allow for the type of browsing seen above along with other advantages.  If you do not, then that will not matter.

#### Variable Descriptions
#####geonum: 
numeric, for merging with data from SDO PostgreSQL Census databases.

#####countyfips: 
Federal Information Processing Standards (FIPS) Number for each County, without leading zeros.

#####year: 
Estimate year, specifically reflects July 1 of each year.

#####jobs: 
Estimates number of jobs in that industry for that county and year.  

#####total_flag
T for an industry total, S for a sub-industry.

#####sector_name:
Name of sector that the estimate was made for.

#####sector_id: 
Numeric ID for the sector or industry the estimate was made for.

Here is a listing of those included and their definition:
```{r, echo=FALSE}
print(county_jobs%>%filter(countyfips==31, year==2013)%>% select(sector_id, sector_name))%>% print(n=100)
```


